Nextstrain: real-time tracking of pathogen evolution

### 2018-04-17 replyToReviewer.tex

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}

We thank the reviewer for their kind words, and have addressed all of the points (detailed below).

@@

\begin{quote}
1. Page 1, Line 39: "ongoing and ever-present" is somewhat redundant; I'd just go with "ever-present"

2. Page 1, Line 40: perhaps change "recent epidemics" to "recent events" so "epidemics" isn't repeated 3x in the sentence.
\end{quote}

We have changed the wording of the manuscript to address these points.

\begin{quote}
3. Page 1, Line 42: it might be helpful to point the reader to a general reference on phylodynamics, assuming the reference limit hasn't already been reached.
\end{quote}
We have reworded this section to include a reference to ``Viral Phylodynamics'', Voltz \textit{et al.} 2013.

\begin{quote}
4. Page 1, Line 43 (column 2): could references to the three databases (papers or URLs) be included here?
\end{quote}
We have added URLs to the text.

\begin{quote}
5. Page 2, Line 35: if it can be explained in a very short sentence, it might be interesting to mention the caveat of "nearly all outbreaks" - which ones aren't amenable to Nextstrain-ification?
\end{quote}

\begin{quote}
6. Page 2, Line 51: it should be noted that the transmission inference depends on how well the outbreak in question was sampled - if data is missing, important transmission links could be missed.
\end{quote}

\begin{quote}
7. Page 2, Line 20 (column 20): is there an example of one of these mutations that could be mentioned here?
\end{quote}

\begin{quote}
There is a fair amount of passive voice in the manuscript - popping the text into readable.io to find these instances and changing the construction to make them active will make the writing a little more lively (and it usually frees up some words in the word count).

\end{quote}

### 2018-04-17 whitepaper.tex
reviewer points 3 & 4

Nextstrain consists of three components:
\textit{Fauna} is a collection of Python scripts to maintain a database representative of all available sequences and related metadata, sourced from public repositories such as NCBI, GISAID and ViPR, as well as GitHub repositories and other sources of genomic data.
\textit{Augur} is the bioinformatics pipeline, performing subsampling, alignment, phylogenetic inference, temporal dating of ancestral nodes and discrete trait geographic reconstruction, including inference of the most likely transmission events.
This leverages the maximum likelihood phylodynamic analyses implemented in TreeTime \citep{sagulenko2017treetime}, allowing a full analysis of the entire Ebola epidemic ($n$=1581 genomes) in under 2 hours on a modern laptop computer.
These scripts separate generic core functionality from a light pathogen-specific layer such that they are easily co-opted to different pathogens.
Outputs from the pipeline are uploaded online and immediately available through the visualisation software \textit{Auspice}, located at \href{http://nextstrain.org}{nextstrain.org}.
This approach is similar in concept to Nextflu \citep{neher2015nextflu} however extended and generalised to different viral pathogens.
There is a growing need for surveillance of non-influenza viruses \citep{tang2017global}, and Nextstrain is able to be extended to nearly all outbreaks with readily accessible genomic data.

>>>>

Nextstrain consists of three components:
\textit{Fauna} is a collection of Python scripts to maintain a database representative of all available sequences and related metadata, sourced from public repositories such as NCBI (www.ncbi.nlm.nih.gov), GISAID (www.gisaid.org) and ViPR (www.viprbrc.org), as well as GitHub repositories and other sources of genomic data.
\textit{Augur} is a suite of tools to perform phylodynamic analysis \citep{voltz2013}, including subsampling, alignment, phylogenetic inference, temporal dating of ancestral nodes and discrete trait geographic reconstruction, including inference of the most likely transmission events.
This leverages the maximum likelihood analyses implemented in TreeTime \citep{sagulenko2017treetime}, allowing a full analysis of the entire Ebola epidemic ($n$=1581 genomes) in under 2 hours on a modern laptop.
These scripts separate generic core functionality from a light pathogen-specific layer such that they are easily co-opted to different pathogens.
Outputs from the pipeline are uploaded online and immediately available through the visualisation software \textit{Auspice}, located at \href{http://nextstrain.org}{nextstrain.org}.
This approach is similar in concept to Nextflu \citep{neher2015nextflu} however extended and generalised to different viral pathogens.
There is a growing need for surveillance of non-influenza viruses \citep{tang2017global}, and Nextstrain is able to be extended to nearly all outbreaks with readily accessible genomic data.
